---
title: "Test constructions and test scores"
author: "Jinnie shin | jinnie.shin@coe.ufl.edu"
output: html_notebook
---

```{r setup, include=FALSE}
library(psych)
library(ggplot2)
library(CTT)

```

## 1. Exam Data (MC-items)
A data frame with 100 observations on the following 20 items. 

```{r, source, exercise=TRUE, echo=FALSE}
data(CTTdata)
attach(CTTdata) 
head(CTTdata)
data(CTTkey)

data <- score(CTTdata,CTTkey, output.scored=TRUE)
response <- as.data.frame(data$scored)
head(response) # this saves the item score (0 or 1)
```

## 2. Item Difficulty 
- The proportion of those responding correctly

```{r,  data, exercise=TRUE, echo=FALSE, exercise.setup="source"}

# Let's calculate the difficulty of item 2

p = sum(response$i2)/100
print(p)

```

Let's calculate the difficulty of item 10
```{r,  data2, exercise=TRUE, echo=FALSE, exercise.completion=TRUE, exercise.setup="source"}

# Let's calculate the difficulty of item 10



```

Let's calculate the variance of item 2
```{r,  data333, exercise=TRUE, echo=FALSE, exercise.completion=TRUE, exercise.setup="source"}

# Let's calculate the variance of item 2
p = sum(response$i2)/100
print(p)

# Let's look at the number of correct and incorrect responses




```

Let's calculate the variance of item 10
```{r,  data3335, exercise=TRUE, echo=FALSE, exercise.completion=TRUE, exercise.setup="source"}

####### Your Code here ################




######################################
```

## 3. Inter-item Reliability 
- inter-item correlations from joint frequency distribution tables

```{r,  data3, exercise=TRUE, echo=FALSE, exercise.setup="source"}
# Let's compute the inter-item correlation score between item 7 and 9 
# First, we will have to create a joint frequency distribution table 

table(response$i7, response$i9)

```

```{r, data4, exercise=TRUE, echo=FALSE, exercise.lines=15, exercise.setup="source"}

# Second, we will have to identify the following components (p16, Topic 2)

pjk = 
  
pj = 
  
pk = 
  
qj = 
  
qk = 

# The final outcome will be: 
rho = (pjk-(pj*pk))/(sqrt(pj*qj*pk*qk))  
```

Let's create a function to calculate this more efficiently
```{r, data44, exercise=TRUE, echo=FALSE, exercise.lines=15, exercise.setup="source"}

iic <- function(r1, r2){ # the function takes r1 (response1) and r2 (response2) as input
       jd = table(r1,r2) #joint frequency distribution 
       ############# YOUR CODE HERE ##########
       
       
       
       #######################################
}

# Let's test this out!

```

## 4. Total Test scores 
Total test scores are frequently used in low stakes assessment

```{r, data5, exercise=TRUE, echo=FALSE, exercise.lines=10, exercise.setup="source"}
# total score can be computed using rowSums()

rowSums(response)[13]

###### YOUR CODE HERE #########


################################
```

Let's compute the average test score and compare it with the total item difficulty 
```{r, data6, exercise=TRUE, echo=FALSE, exercise.lines=16, exercise.setup="source"}
# 1. Average Test Score 
###### YOUR CODE HERE #########
ave_test_score = 
################################

# 2. Sum of all item difficulty
###### YOUR CODE HERE #########
sum_dif = 
################################

# 3. Check whether ave_test_score == sum_dif 
###### YOUR CODE HERE #########



```

## 5. Exercise 

#### 1. Most difficult and least difficult items 
```{r, ex1, exercise=TRUE, echo=FALSE, exercise.lines=6, exercise.setup="source"}




```

#### 2. Item with the highest variance and lowest variance 

```{r, ex2, exercise=TRUE, echo=FALSE, exercise.lines=8, exercise.setup="source"}




```

#### 3. Inter-item correlation between the items from Q1
```{r, ex3, exercise=TRUE, echo=FALSE, exercise.lines=15, exercise.setup="source"}




```

#### 4. Inter-item correlation between the items from Q2
```{r, ex4, exercise=TRUE, echo=FALSE, exercise.lines=15, exercise.setup="source"}




```

#### 5. Total score histogram 
```{r, ex5, exercise=TRUE, echo=FALSE, exercise.lines=8, exercise.setup="source"}




```

#### 5. Total score to z-score and histogram 
```{r, ex6, exercise=TRUE, echo=FALSE, exercise.lines=5, exercise.setup="source"}





```
