---
title: "Week 5 - Reliability"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library (readr)
```

## Cronbach's Alpha 

$$ \alpha = \frac{k}{k-1}\left(\frac{2\sum_{i\neq{j}}\sigma_{ij}}{\sigma_{c}^2}\right)\text{,} $$ 

```{r, ca, echo=TRUE, exercise=TRUE, exercise.lines= 23}
urlfile="https://raw.githubusercontent.com/jinnieshinufl/EDF-6436/main/lab/LabR6436/inst/tutorials/Week5/data/Cronbach_data.csv"
data<-read_csv(url(urlfile), col_names=TRUE) 
data<- as.data.frame(data)

# calculating total summated scores
X = rowSums(data)

# calculating variance of total summated scores
varX = var(X)

# calculating variance of each item score
var_matrix = var(data)
vars = diag(var_matrix)

# calculating sum of item variances
sum_vars = sum(vars)

# obtaining number of items
n_items = length(vars)

# calculating alpha
alpha = (n_items/(n_items-1))*(1-(sum_vars/varX))

print(alpha)
```

#### Using the `CTT()` package fromt the last lection
```{r, ctt2, exercise=TRUE}

library(CTT)
reliability(data, itemal=TRUE)

itemAnalysis(data)$itemReport
```

#### Reliability and Confidence Interval 

If we know the mean and variance of the observed scores on a test, as well as the reported reliability, we can estimate a confidence interval around an estimated true score. Let’s use Alba’s predicted true score.


```{r ci, exercise=TRUE, echo=FALSE, exercise.lines=30}
rel <- 
SEM <- 
VarX <- 
Var_E  <- (1-rel)*VarX
SE_Msr <- sqrt(Var_E) 
LB     <- x - 1.96*SEM
UB     <- x + 1.96*SEM

``` 

## Spearman-Brown Prophecy

The Spearman-Brown prophecy formula is great for seeing how many more (or fewer) items we will need to achieve a desired reliability.

$$ \rho_{_{CC'new}} = \frac{k\rho_{_{CC'old}}}{1 + (k - 1)\rho_{_{CC'old}} } $$  

For instance, how many more items do we need to achieve the reliability of 0.90 

$$ k = \frac{\rho_{_{CC'new}}(1 - \rho_{_{CC'old}})}{\rho_{_{CC'old}}(1 - \rho_{_{CC'new}}) }\\ $$ 

```{r, sp, exercise=TRUE}
alpha = 0.83
k = (0.90*(1-alpha))/(alpha*(1-0.90))

library(CTT)
relachieved_if_doubled <- spearman.brown(alpha, input = 1.8, n.or.r = "n")
relachieved_if_doubled
``` 
