---
title: "Item Response Theory"
author: "Jinnie shin | jinnie.shin@coe.ufl.edu"
date: "03/29/2022"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
#install.packages("ltm")
#install.packages("eRm")
library(ltm) 
library(eRm)
library(learnr)
library(questionr)
library(corrplot)
```

## 1. Data 

We will call the data (`Rasch_data.txt`) which includes 21 items answered by 200 students. 

```{r, data, exercise=TRUE, echo=TRUE}
data <- read.table("./data/Rasch_data.txt",header=FALSE,sep="")
head(data)
#Change the text file into matrix format that ltm can read
data <- as.matrix(data[1:200,1:20])
head(data)
```

## 2. Rasch Model 

we will run a rasch model using a package called `ltm` and the function called `rasch()`

### Running a rasch model 
```{r, rasch, exercise=TRUE, echo=FALSE, exercise.setup = "data", exercise.lines=12}

#Running rasch model 
# the constraints are needed to hold all discrimination parameters to <?>

RaschModel <- rasch(data, # takes your data as your first input 
                    constraint= # requires a constraint 
                    )

# Let's take a look at the summary of the model results 


```

- Help!
```{r, help, exercise=TRUE}

```

### Extracting estimated parameters 
- Which parameters are we freely estimating in this model?

```{r param, exercise=TRUE, echo=FALSE,  exercise.setup = "rasch", exercise.lines=10}
#Extracting estimated parameters

#`coef()`` function let you look at the estimated parameters and we will call this `item.pars`

# Let's present it as a data frame using `as.data.frame()`: default r function 


```

### Getting ability level 
```{r ability, exercise=TRUE, echo=FALSE,  exercise.setup = "param", exercise.lines=14}

#Getting ability estimates associated with each response vector
# we need to use a new argument `resp.patterns = <data name>`
fscores <- 

#(i.e., thetas are z values)
theta = fscores$score.dat$z1

assign("theta", theta, envir=globalenv())   
```

#### Compare the ability level estimate with the total score 

```{r, newdataset, exercise=TRUE, exercise.setup = "data", exercise.lines=11}
# step 1: let's compute the total score 


# step 2: extract lines (rows) with the same total score  == 7


# step 3: compare with the factor score 

```

### Item Characteristic Curve 

$$ p = \frac{e^{1.7*(theta - b)}}{1 + e^{1.7*(theta - b)}} $$ 
# Let's create a function that computes the probability given the theta and 

```{r, icc-value, exercise=TRUE, echo=FALSE, exercise.lines=25}
# STEP 1
# this function takes two input values -> theta and b-parameter 
prob <- function(theta, b){
   # Provide the formula here 
  
}
# STEP 2
# Let's test it quickly ==> if b=2, theta=2 prob=?



# STEP 3
# Let's create ICC for an item with b-parm =2.0
# ability range from -4 to 4 (increase by +1)

# STEP 4: Uncomment the last three lines when you are ready :-) 
#plot(c(p0, p1, p2, p3, p4, p5, p6, p7, p8), 
#     x = c(-4, -3, -2, -1, 0, 1, 2, 3, 4), 
#     xlab = "Ability", ylab="Probability")

```

### Plotting 
- Here's your shortcut! `plot()`

```{r plot, exercise=TRUE, echo=FALSE , exercise.setup = "rasch"}
#Plotting ICCs for all items
plot(RaschModel, main="ICC from Rasch Model")

```


#### Plotting for a single item 

```{r plot_one, exercise=TRUE, echo=FALSE,  exercise.setup = "rasch"}

#Plotting ICC for a single item
plot(RaschModel, items = c(2), main="ICC of Item 2")

```

### Item Information Function 
$$p = \frac{\exp^{1.7*(\theta - b)}}{1 + \exp^{1.7*(\theta - b)}}$$
$$ I(\theta) = p(1-p) $$

```{r, iif, exercise=TRUE, echo=FALSE, exercise.lines=30}

# STEP 1
# this function takes two input values -> theta and b-parameter 
info <- function(theta, b){
  # provide the equation here 
}

# STEP 2
# Let's test it quickly ==> if b=2, theta=2 info=?

# STEP 3
# Let's create ICC for an item with b-parm =2.0
# ability range from -4 to 4 (increase by +1)


# STEP 4: Uncomment the last three lines when you are ready :-) 
plot(c(info0, info1, info2, info3, info4, info5, info6, info7),
     x = c(-4, -3, -2, -1, 0, 1, 2, 3),
     xlab = "Ability", ylab="Information")
```

### Plotting Item and Test Information Curve (TCC)

```{r icc2, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
#Plotting TIC
plot(RaschModel,  type=c("IIC"), main="Item Information Function - All")

#Plotting ICC
plot(RaschModel, items = c(2), type=c("IIC"), main="Item Information Function - Item 2")

```

```{r question, echo=FALSE}

question('What is the information for this item [b=0.7694] at θ = 0.7694?', 
  answer("0.25", correct=TRUE), 
  answer("0.025"), 
  answer("0.0025"),
  answer("0.00025"))
```

### Evaluating the Model 
#### 1. Test for the item fit using `item.fit`
```{r 1pleval, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
item.fit(RaschModel,simulate.p.value=T) # are there any particular items with small p-values? 


```
#### 1. Test for the person fit using `person.fit`
```{r 1pleval2, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
person.fit(RaschModel,simulate.p.value=T) 

```


#### 3. Estimate ability score (theta) & Person-item map
> Person-item map: Displays the location of person abilities and item difficulties respectively along the same latent dimension. 

```{r 1pleval3, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
fscores = factor.scores(RaschModel) # don't provide the response matrix thist ime 
plot(fscores)
plot(fscores, include.items = TRUE, main ="Item Person Map under 1PL model")
```


#### 4. Test for Unidimensionality (this will run for a long time)
```{r 1pleval4, exercise=TRUE, echo=FALSE, exercise.setup = "rasch"}
unidimTest(RaschModel, data)
# the null hypothesis = unidimensional
# if the test is significant =/ unidimensional 
```

## 2. 2-PL and 3-PL 

### Data

Let's call the data (`2&3PL_data.txt`) which included a total of 3,000 students' responses on 20 dichotomously scored items. 

```{r, data2, exercise=TRUE, echo=FALSE}

temp <- read.table("./data/2&3PL_data.txt",header=FALSE,sep="")

#Change the text file into matrix format that ltm can read
data <- as.matrix(temp[1:3000,1:20])

```

### Running 2PL Model

```{r, 2pl, exercise=TRUE, echo=FALSE, exercise.setup = "data2", exercise.lines=30}

#Running 2-PL model
two.pl.model <- ltm(data ~ z1)

#Assessing item fit
#null hypothesis is that observed data matches expected values
#item.fit(two.pl.model,simulate.p.value=TRUE)

#Extracting estimated parameters
item.pars.two.pl <- coef(two.pl.model)
item.pars.two.pl <- as.data.frame(item.pars.two.pl)
item.pars.two.pl



#Plotting ICCs 
plot(two.pl.model, main="ICC Curves under 2-PL Model")

#Creating an item person map
fscores <- factor.scores(two.pl.model)
plot(fscores, include.items = TRUE, main ="Item Person Map under 2PL model")

assign("two.pl.model",two.pl.model, envir=globalenv())
```

### Running 3PL Model

```{r, 3pl, exercise=TRUE, echo=FALSE, exercise.setup = "data2"}

#Running 3PL model 
three.pl.model <- tpm(data)

#Assessing item fit
#null hypothesis is that observed data matches expected values
#item.fit(three.pl.model, simulate.p.value=TRUE)

#Extracting estimated parameters
item.pars.three.pl <- coef(three.pl.model)
item.pars.three.pl <- as.data.frame(item.pars.three.pl)
item.pars.three.pl

#Plotting ICCs
plot(three.pl.model, main="ICC Curves under 3-PL Model")

#Creating an item person map
fscores <- factor.scores(three.pl.model)
plot(fscores, include.items = TRUE, main ="Item Person Map under 3PL model")
assign("three.pl.modell",three.pl.model, envir=globalenv())
```

## 3. Model-fit 1PL (eRm-other pacakges)

```{r, rasch2, exercise=TRUE, echo=FALSE, exercise.setup = "data", exercise.lines=30}

library(eRm) # loading a library 
one.pl.model <- RM(data) #using a `RM()` function to fit a rasch model using our dataset
summary(one.pl.model ) # this will give you a summary of the outcome
betas <- -coef(one.pl.model ) #we are extracting only the difficulty parameter, but reverse code them
plotjointICC(one.pl.model , cex = .6) # we can plot ICC 
plotPImap(one.pl.model , cex.gen = .55) # plot person-item map 

person.fit = person.parameter(one.pl.model)
itemfit(person.fit)
plotPWmap(one.pl.model)

plotPWmap(one.pl.model, pmap=TRUE, imap=FALSE)
assign("one.pl.model",one.pl.model, envir=globalenv())
```

## 4. Model-fit Comparisons (eRm-other pacakges)
```{r, rasch3, exercise=TRUE, echo=FALSE, exercise.lines=5}
anova(two.pl.model, three.pl.model)
```

```{r, rasch4, exercise=TRUE, echo=FALSE, exercise.lines=5}
cor(coef(two.pl.model)[, 1], coef(three.pl.model)[, 2])
```
