---
title: "Explaratory Factor Analysis"
author: "Jinnie shin | jinnie.shin@coe.ufl.edu"
date: "3/5/2024"
output: html_document
---

```{r setup, include=FALSE}
library(psych)
library(EFAtools)
library(readr)
```

## 1. Cognition Data 
The `cognition` data set includes measures of vocabulary (vocab), reading comprehension (reading), sentence completion (sentcomp), mathematics (mathmtcs), geometry (geometry), and analytical reasoning (analyrea).

```{r source}
urlfile="https://raw.githubusercontent.com/jinnieshinufl/EDF6436-Spring2024-/main/interest.csv"
interest<-read_csv(url(urlfile), col_names=TRUE) 
interest<- as.data.frame(interest)

cognition <- subset(interest, select = vocab:analyrea)
head(cognition)
assign("cognition", cognition, envir=globalenv()) 

```

### Descriptive statistics of the data
Looking at the descriptive statistics, it appears that the variables may have been standardized (indicated by the means of the variables being ap- proximately zero and the variances being approximately one). 

```{r, score_data, exercise=TRUE, echo=TRUE}
# Descriptive analysis - summary table 

```

### Assumption checking 1 - Normality 
We should also check whether the variables are normally distributed 
```{r, normal, echo=FALSE, exercise=TRUE, exercise.lines=13}
# Visually inspect whether the distribution is bell shaped.
### 1. histogram (column='vocab')





### 2. density plot (column = 'vocab')

```

#### Q-Q Plot 
```{r, normal2, echo=FALSE, exercise=TRUE, exercise.lines=5}
### 3. qq-plot (column='vocab')


```

#### Shapiro-Wilk Test 
H0 = the population is distributed normally.
```{r, norm3, exercise=TRUE, echo=FALSE ,exercise.lines=5}
### 4. Shapiro-Wilk Test  (column='vocab')

#shapiro.test()
```

## 2. Assumption Checking 
The first step in an EFA procedure is to test whether your data is suitable for factor analysis. To this end, the **EFAtools package** provides the ``BARTLETT()`` and the ``KMO()`` functions. The Bartlett’s test of sphericity tests whether a correlation matrix is significantly different from an identity matrix (a correlation matrix with zero correlations between all variables). This test should thus be significant. The Kaiser-Meyer-Olkin criterion (KMO) represents the degree to which each observed variable is predicted by the other variables in the dataset and thus is another indicator for how correlated the different variables are.

```{r, assumption, echo=FALSE, exercise=TRUE}

# Assumption 1: Data Sphericity - Bartlett's test of sphericity
BARTLETT(cognition) # from EFATools 

# Assumption 2: Sampling Aqequacy - The Kaiser-Meyer-Olkin measure 
KMO(cognition) # from EFATools 

```

## 3. Correlation 
```{r, corr, echo=FALSE, exercise=TRUE}
correlations <- cor(cognition) 
round(correlations, 3)

library(corrplot) #correlation plot 
corrplot(correlations, method="number",type="lower")
assign("correlations", correlations, envir=globalenv()) 
```

## 4. Determining the number of factors 
```{r, corr3, echo=FALSE, exercise=TRUE}

#### 1. K1 Method
# Let's do eigenvalue decomposition using the function `eigen()` => default 
eigen_decomp <- 

print(eigen_decomp)
```

```{r,scree, echo=FALSE, exercise=TRUE }
### 2. Scree plot (elbow method)

scree(cognition) #from the psych package

```

```{r, para, echo=FALSE, exercise=TRUE}
### 3. Parallel analysis 
fa.parallel(cognition, fm = "ml") #from the psych package 
```

## 5. Factor Analysis - continuous 
Using the `fa()` function from the `psych` package, we can now specify a two factor solution. We will use the commonly adopted `promax` rotation (oblique) as factors tend to correlated most cases.

```{text}
"Factoring method fm="minres" will do a minimum residual as will fm="uls". Both of these use a first derivative. fm="ols" differs very slightly from "minres" in that it minimizes the entire residual matrix using an OLS procedure but uses the empirical first derivative. This will be slower. fm="wls" will do a weighted least squares (WLS) solution, fm="gls" does a generalized weighted least squares (GLS), fm="pa" will do the principal factor solution, fm="ml" will do a maximum likelihood factor analysis. fm="minchi" will minimize the sample size weighted chi square when treating pairwise correlations with different number of subjects per pair. fm ="minrank" will do a minimum rank factor analysis. "old.min" will do minimal residual the way it was done prior to April, 2017 (see discussion below). fm="alpha" will do alpha factor analysis as described in Kaiser and Coffey (1965)"
```

```{r, help, echo=TRUE, exercise=TRUE}

```

```{r, corr2, echo=TRUE, exercise=TRUE, exercise.lines=13}
# Factor analysis using the psych package -> `fa()`
one_factor <- fa(cognition,  # data
                 nfactors = 1, # number of factors 
                 fm = "pa",  # method = 
                 rotate = "promax") # rotation 


one_factor # call the results 


assign("one_factor", one_factor, envir=globalenv()) # this is just to connect the chunks
```

We can prepare this output so it is more readable. For instance, we can ask that the loadings that do not exceed a cut-off value, such as .32, be excluded from the printout.

```{r, result, exercise=TRUE, echo=FALSE}
print(one_factor, cut = .32, sort = TRUE, digits = 3)
```

### Diagram 
```{r,eval, exercise=TRUE, echo=FALSE}
fa.diagram(one_factor, digits = 2, main = "Factor Diagram", 
           cut = .32, 
           simple = F, 
           errors = T)

```

### Factor score 
```{r,fscores, exercise=TRUE, echo=FALSE}
#factor.scores()

```

### Exercise 
Let's try to run the two factor model 
```{r, two_factor, exercise=TRUE, echo=FALSE, exercise.lines= 15}
# run the two factor model and call it `two_factor` 
# hint: a more flexible rotation? 




# print the two factor model results with the cut-off score of .30

```

- Do you think we should change the rotation method. Why or Why not? 

- Interpret the model-fit evaluation results. 

## 6. Factor Analysis - ordinal
We could sometimes have ordinal data collected (e.g., rating scale). We will use the Humor Style Questionnaire dataset which includes responses from 1071 participants on 32 questions. 

|     | Questionnaire                                                                                                              |
|-----|----------------------------------------------------------------------------------------------------------------------------|
| Q1  | I usually don't laugh or joke around much with other people.                                                               |
| Q2  | If I am feeling depressed, I can usually cheer myself up with humor.                                                       |
| Q3  | If someone makes a mistake, I will often tease them about it.                                                              |
| Q4  | I let people laugh at me or make fun at my expense more than I should.                                                     |
| Q5  | I don't have to work very hard at making other people laugh. I seem to be a naturally humorous person.                      |
| Q6  | Even when I'm by myself, I'm often amused by the absurdities of life                                                       |
| Q7  | People are never offended or hurt by my sense of humor.                                                                    |
| Q8  | I will often get carried away in putting myself down if it makes my family or friends laugh.                               |
| Q9  | I rarely make other people laugh by telling funny stories about myself.                                                    |
| Q10 | If I am feeling upset or unhappy I usually try to think of something funny about the situation to make myself feel better. |
|     | ...                                                                                                                        |
| Q32 | Letting others laugh at me is my way of keeping my friends and family in good spirits.    

```{r}
url <- "https://assets.datacamp.com/production/course_4249/datasets/humor_dataset.csv"
data <- read.csv(url, sep = ";")
assign("data", data, envir=globalenv())  # This is just to connect the chunks 
```


```{r, gl, exercise=TRUE, echo=FALSE, exercise.lines=9}
data <- data.frame(lapply(data, ordered)) 

# data overview 

# summary statistics


```

### Data cleaning 
Anything you noticed? 
```{r clean, exercise=TRUE, exercise.lines=8}
# recoding 



```
#### Correlation Matrix 

For the factor analysis, because our data are ordered categorical and not continuous, we should use the a polychoric matrix instead of the raw values. 
We can directly specify this in the `fa()` function using the `cor = "poly"` argument.

```{r, poly, exercise=TRUE, echo=FALSE}
temp   <- fa(data[,1:32], 
            fm       = "pa", 
            rotate   = "promax",
            cor      = "poly")

R_poly <- temp$r # R = polychoric matrix 

assign("R_poly", R_poly, envir=globalenv())
```

#### Assumption checking 
```{r, assumption2, exercise=TRUE, echo=FALSE}
KMO(R_poly)

cortest.bartlett(R_poly, n= nrow(data))

```

#### Determining the number of factors 
```{r, cor55, echo=FALSE, exercise=TRUE}

#### 1. K1 Method

eigen_decomp <- eigen(R_poly) 
round(eigen_decomp$values, 3)

### 2. Scree plot (elbow method)

scree(R_poly)

### 3. Parallel analysis 
fa.parallel(R_poly, fm = "ml")
```

#### Factor Analysis on the Polychoric Matrix
```{r, fa2, exercise=TRUE, echo=FALSE}
fafit <- fa(R_poly, 
            nfactors = 4, 
            fm       = "pa", 
            rotate   = "promax")
print(fafit, cut = .32, sort = TRUE, digits = 3)

assign("fafit", fafit, envir=globalenv())
```

#### Pattern Matrix - Factor loading 
```{r, pattern, exercise=TRUE, echo=FALSE}
PattM   <- fafit$loadings[1:4,]

print(PattM)
```

#### Factor Score 
When we have a raw data set, for example with each row representing a person, we can estimate those persons’ factor scores.
```{r fs, exercise=TRUE, echo=FALSE}
Factor_scores <- factor.scores(data[,1:32], fafit)
Factor_scores 
``` 

### Factor Digram 
```{r da2, exercise=TRUE, echo=FALSE}
fa.diagram(fafit, digits = 2, main = "Factor Diagram", 
           cut = .32, 
           simple = F, 
           errors = F)
```

### Naming Factors 
- Let's go back to our survey question and try to name the factors
```{text}
Factor 1:
  
Q25. I don’t often joke around with my friends.
Q17. I usually don’t like to tell jokes or amuse people
Q1. I usually don’t laugh or joke around much with other people.
Q13. I laugh and joke a lot with my closest friends.
Q21. I enjoy making people laugh
Q29. I usually can’t think of witty things to say when I’m with other people
Q5. I don’t have to work very hard at making other people laugh. I seem to be a naturally humorous person
Q9. I rarely make other people laugh by telling funny stories about myself
```

```{text}
Factor 2: 
  
Q18. If I’m by myself and I’m feeling unhappy, I make an effort to think of something funny to cheer myself up
Q10. If I am feeling upset or unhappy I usually try to think of something funny about the situation to make myself feel better.
Q26. It is my experience that thinking about some amusing aspect of a situation is often a very effective way of coping with problems.
Q14. My humorous outlook on life keeps me from getting overly upset or depressed about things.
Q2. If I am feeling depressed, I can usually cheer myself up with humor
Q30. I don’t need to be with other people to feel amused 
Q6. Even when I’m by myself, I’m often amused by the absurdities of life.
Q22. If I am feeling sad or upset, I usually lose my sense of humor.
```
 
```{text}
Factor 3: 
  
Q20. I often go overboard in putting myself down when I am making jokes or trying to be funny.
Q8. I will often get carried away in putting myself down if it makes my family or friends laugh.
Q32. Letting others laugh at me is my way of keeping my friends and family in good spirits.
Q12. I often try to make people like or accept me more by saying something funny about my own weaknesses, blunders, or faults.
Q14. My humorous outlook on life keeps me from getting overly upset or depressed about things.
Q2. If I am feeling depressed, I can usually cheer myself up with humor.
Q30. I don’t need to be with other people to feel amused 
Q6. Even when I’m by myself, I’m often amused by the absurdities of life.
Q22. If I am feeling sad or upset, I usually lose my sense of humor.
 
```

```{text}
Factor 4: 
  
Q31. Even if something is really funny to me, I will not laugh or joke about it if someone will be offended.
Q15. I do not like it when people use humor as a way of criticizing or putting someone down
Q7. People are never offended or hurt by my sense of humor.
Q11. When telling jokes or saying funny things, I am usually not very concerned about how other people are taking it.
Q27. If I don’t like someone, I often use humor or teasing to put them down.
Q3. If someone makes a mistake, I will often tease them about it.
Q23. I never participate in laughing at others even if all my friends are doing it.
Q19. Sometimes I think of something that is so funny that I can’t stop myself from saying it, even if it is not appropriate for the situation.
 
```

## 7. Naming Factors (continue)

- https://en.wikipedia.org/wiki/Humor_styles 
- https://www.psytoolkit.org/survey-library/humor-hsq.html

#### Affiliative humor
Affiliative humor is defined as the style of humor used to enhance one's relationships with others in a benevolent, positive manner. This style of humor is typically used in a benevolent, self-accepting way. Individuals high in this dimension often use humor as a way to charm and amuse others, ease tension among others, and improve relationships. 

#### Self-enhancing humor
Self-enhancing humor is a style of humor related to having a good-natured attitude toward life, having the ability to laugh at yourself, your circumstances and the idiosyncrasies of life in constructive, non-detrimental manner. It is used by individuals to enhance the self in a benevolent, positive manne

#### Aggressive humor
Aggressive humor is a style of humor that is potentially detrimental towards others. This type of humor is characterized by the use of sarcasm, put-downs, teasing, criticism, ridicule, and other types of humor used at the expense of others. 

#### Self-defeating humor
Self-defeating humor is the style of humor characterized by the use of potentially detrimental humor towards the self in order to gain approval from others. Individuals high in this dimension engage in self-disparaging humor in which laughter is often at their own expense. 
